Structuring College Access

### 2025-04-04 index.qmd
adding background and lit section

### Introduction {.hide}

@@

<!-- Research interest: how quantification reproduces social stratification -->

@@

# Background and Theory

@@

- Meant to reproduce the recruiting territories of admissions officers
- @RN4982 has 143 Geomarkets in 3 regions; EPS has 304

>>>>

### [__College Board _Student Search Service_, order \#448922__](./assets/images/geomarket_orders_redacted.pdf){target="_blank"} {.hide}

__Sorting students into colleges, perspectives from sociology:__

- Status attainment model [e.g., @RN915]
- Cultural capital model [e.g., @RN1228]
<!-- - "Schools as sorting machines" [@RN4873] -->
- Enrollment management behavior of colleges [e.g., @RN3519]

__Sociology misses an important mechanism of social reproduction:__

- Third-party vendors/products sort students on behalf of colleges
- Recommendations based on historical snapshots of student demand

__"Geomarkets" are an input in two College Board recruiting products:__

- Student Search Service (est. 1972); Enrollment Planning Service (est. 1984)

__Research questions:__

### Introduction og {.hide}

@@

 -->

@@

### Presentation Outline {.hide}

1. Enrollment management: Background \& scholarship
1. Empirical case and theory
1. Methods
1. Results
1. Discussion
    - Creating a "knowledge infrastructure" [@RN5025] to study the role of private firms and private equity funding in education

# Enrollment Management: Background \& Scholarship

### __Enrollment Management (EM)__ {.hide .columns-2}

::: columns

:::: {.column width=65%}

__Enrollment management__

- EM profession integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]
- EM office controls recruiting, admissions, fin aid

__Interventions along enrollment funnel__

- Convert prospects to leads 
  - Purchase student lists
- Convert leads/inquiries to applicants
  - Email, mail, school visits
- Convert applicants to admits
  - Holistic review [e.g., @RN4146]
- Convert admits to enrollees
  - Financial aid offers [e.g., @RN2241]
::::

:::: {.column width=35%}

__Enrollment Funnel__
```{r, echo=FALSE, out.width="85%", fig.cap=""}
knitr::include_graphics("./assets/images/enroll_funnelv2.png")
```
::::

:::

### __Scholarship on Recruiting "in the Wild"__{.hide}

__Sociological ethnographies of enrollment management__

- Private Colleges: visit affluent "feeder schools" to maintain relationships with HS guidance counselors [@RN3519; @RN4407]
- For-profits: door-to-door visits to local, low-income communities of color [@cottom2017lower]

__Web-scraped the travel schedules of public research and selective private univs__

- Visit rich, mostly White public schools [@RN4758]
- Disproportionate visits to private schools [@RN5023]
- "Recruitment redlining": circuitous avoidance of Black and Latinx communities along visit paths [@RN4759] [INSERT CLICKABLE IMAGE OF LA MAP]

___Recruiting not a practice wholly done by colleges...___

- Data vendors and consultancies help colleges find students

### <span style="font-size:0.9em;">__U.S. Higher Education Market: A National Voucher System__</span> {.hide}

__Federal government__

- Takes a "steer, don't row" approach
- Subsidizes higher education by providing Title IV grants and loans to students
- Department of Education regulates Title IV institutions

__Students__

- Attend college to improve career, social, and civic outcomes
- Pay for college using: 
  - Household savings; grants and loans from federal, state, private sources

__Direct providers (Title IV institutions)__

- A mix of public, private nonprofit, for-profit providers
- Funding follows students to whichever institution they enroll in
  - A national voucher system
- Colleges have a financial incentive to enroll students, so competition for students

### __Student Lists are a Match-Making Intermediary__{.hide}

__@RN2247 on emergence of a national higher education system:__

- Information costs were primary barrier to efficient matches
  - Students want to attend best college, but don't know options or costs
  - Colleges want to enroll best students, but don't know who or where they are
- SAT causes "dramatic fall" in cost of "information about students" (p. 102)

__"Lead generation" is process of connecting merchants to potential consumers__

- "List-based" lead generation based on direct-mail model (legacy)
  - In 1972, College Board began selling lists of prospects to colleges
- "Behavioral-based" targeting (contemporary)
  - Target users of a platform while they use that platform or a partner platform

__Students who opt into College Board _Search_ are [@RN4739]:__

  - 25% more likely to enroll in a four-year university after high school
  - 31% more likely to graduate college within four years 
- Magnitude of effects largest for Black, Latinx, and first-gen college students

### __College Board Lists and Student Outcomes__{.hide}
#### _Howell, Hurwitz, Mabel et al. (2021)_

```{r, fig.align="left"}
knitr::include_graphics("./assets/images/cb-fig-1.png")
```

### __The Student List Project__{.hide}

[THIS SLIDE NEEDS LOTS OF WORK; NEED TO SHOW THE SEARCH FILTERS AS PROBLEMATIC]
Public records requests to public universities in CA, IL, MN, TX, AZ from 2016-2020

- "Order summary" specifying search filter criteria [(Sample)](https://drive.google.com/drive/u/1/folders/18CHMlUIa5OO0zRkET7eSIzYnvk4S9_U1)
- De-identified prospect-level student list [(Sample)](https://drive.google.com/drive/u/1/folders/1rDIAuHLsh3IfglSsqhNuCmYsfnbtc0tV)

__What is the relationship between using academic and geographic filters and the demographic composition of resulting student lists?__ [@RN5012]

- Analyze descriptive relationships observed in list purchases from College Board and HSLS:09 simulated relationships
- Findings: Racial (and socioeconomic) inequity results from:
  - College Board student list databases exclude non-test-takers
  - Test scores and geographic filters compound disparities in excluded prospects

MOTIVATE THIS STUDY

# Emprical Case and Theory

@@

- Geomarkets carve states and metro areas into smaller geographic regions
- Meant to reproduce the recruiting territories of admissions officers
- @RN4982 has 143 Geomarkets in 3 regions; EPS has 304
